A Type System for Safe and Scalable Parallelism using Logical Regions

### 2012-07-11 doc/popl2013/example.tex
Updating the example with the some changes

Listing~\ref{lst:circuit_ex} shows a circuit simulation written (with
one exception) in the core Legion language (see
Section~\ref{sec:legioncore}).  The input is a graph of circuit
elements and connecting wires.  To perform parallel
simulation the graph is partitioned into pieces, after which the
simulation is run for a number of time steps.  Each time step has
three phases: calculation of each wire's current, distribution of charge to
the nodes to which each wire is connected, and updating of each node's voltage.

The top-level function is {\tt simulate\_circuit} (line 15), which is
parameterized on regions {\tt rl}, {\tt rw}, and {\tt rn}.
The wires and nodes are contained in regions {\tt rw} and {\tt rn},
respectively.  The {\tt rl} region holds lists of pointers to the nodes and wires.
When {\tt simulate\_circuit} is invoked it must have both read and
write privileges for the regions {\tt rl}, {\tt rw}, and {\tt rn}
(line 17); the type system enforces that
{\tt simulate\_circuit} only accesses these regions and their
subregions.

Partitioning of the graph is done using {\em colorings} built in the
{\tt color\_circuit} function (line 18-19).  A coloring maps region
elements to ``colors'' (i.e. integers), with each color corresponding
 to a new subregion
to be created.  Note that it is impossible to pick a single static
partitioning scheme that would do well for all graphs.  Dynamic
partitioning using colorings provides the programmer the flexibility
to make partitioning decisions based on input data.

@@

After creating the colorings, the application partitions the node and wire
regions
into {\em subregions} according to the colorings (lines 21-25).  For
simplicity, our example only partitions the circuit into two pieces.
Line 21 uses {\tt pc.1} to partition {\tt rn} into {\tt rn0}, and {\tt rn1}.
Subregion {\tt rn0} will contain all locations with a color of 1 in {\tt pc.1},
while {\tt rn1} will contain all locations with a color of 2.  This
partitioning uses a coloring and results in {\em disjoint} subpartitions.
Constraints are introduced into the static environment describing both the
disjointness of the subregions (i.e. {\tt rn0 $*$ rn1}) and the inclusion of
each subregion in the original region (e.g. {\tt rn0 $\leq$ rn}).  Line
23 is similar, but uses a multicoloring, so the subregions may be {\em aliased},
and only the inclusion constraints are introduced into the static environment.
Although multicolorings are not part of the Core Legion described in 
Section~\ref{sec:legioncore}, they can be easily implemented using a separate 
coloring for each color in the multicoloring.

@@

This function also demonstrates the benefit of being able to dynamically
(re-)discover disjointness.  Safely running both instances of 
{\tt calc\_new\_currents} in parallel depends on knowing that {\tt rw0} and
{\tt rw1} in {\tt execute\_time\_steps} are disjoint.  Similarly, parallel
execution of the instances of {\tt update\_voltage} requires knowing that
{\tt rn0} is disjoint from {\tt rn1}.  In both cases, this knowledge was
statically available in {\tt simulate\_circuit} and could have been captured
in a region relationship, but it would have resulted in much more complicated
code.
Instead, an inexpensive dynamic disjointness check is performed by the Legion
runtime, and the safety of parallel execution follows from the soundness of
the type system.

@@

In addition to privileges, functions can also specify {\em coherence}
on regions.  Coherence specifies what updates the function may
see from other functions using the same region.  If not otherwise
specified, coherence defaults to {\em exclusive}, meaning the
function must appear to execute in program order.  Line 14 shows
an example of {\tt atomic}, a relaxed coherence mode requiring
that operations to {\tt rn} and {\tt rg} must appear atomic relative
to other functions using those regions.  The most relaxed coherence
mode is {\tt simult}; all functions with 
simultaneous coherence are allowed to access the region concurrently
with other functions having simultaneous coherence to the region.

>>>>

Listing~\ref{lst:circuit_ex} shows a circuit simulation written (with
one exception) in the core Legion language (see
Section~\ref{sec:legioncore}).  The input is a graph of circuit
elements and connecting wires.  To perform parallel
simulation the graph is partitioned into pieces, after which the
simulation is run for a number of time steps.  Each time step has
three phases: calculation of each wire's current, distribution of charge to
the nodes connected to each wire, and updating of node voltages.

The entry function is {\tt simulate\_circuit} (line 15), which is
parameterized on regions {\tt rl}, {\tt rw}, and {\tt rn}.
The wires and nodes are contained in regions {\tt rw} and {\tt rn},
respectively.  The {\tt rl} region holds lists of pointers to the nodes and wires.
When {\tt simulate\_circuit} is invoked it must have both read and
write privileges for the regions {\tt rl}, {\tt rw}, and {\tt rn}
(line 17); the type system enforces that
{\tt simulate\_circuit} only accesses these regions and their
subregions at compile-time.

Partitioning of the graph is done using {\em colorings} built in the
{\tt color\_circuit} function (line 18-19).  A coloring maps region
elements to colors (i.e. integers), with each color corresponding
 to a new subregion
to be created.  Note that it is impossible to pick a single static
partitioning scheme that would do well for all graphs.  Dynamic
partitioning using colorings provides the programmer the flexibility
to make partitioning decisions based on input data.

@@

After creating the colorings, the application partitions the node and wire
regions into {\em subregions} according to the colorings (lines 21-25).  For
simplicity, our example only partitions the circuit into two pieces.
Line 21 uses coloring {\tt pc.1} to partition {\tt rn} into {\tt rn0}, and {\tt rn1}.
Subregion {\tt rn0} will contain all locations with a color of 1 in {\tt pc.1},
while {\tt rn1} will contain all locations with a color of 2.  This
partitioning uses a coloring and results in {\em disjoint} subpartitions.
Constraints are introduced into the static environment describing both the
disjointness of the subregions (i.e. {\tt rn0 $*$ rn1}) and the inclusion of
each subregion in the original region (e.g. {\tt rn0 $\leq$ rn}).  Line
23 is similar, but uses a multicoloring, so the subregions may be {\em aliased},
and only the inclusion constraints are introduced into the static environment.
Although multicolorings are not part of the Core Legion described in 
Section~\ref{sec:legioncore}, they can be easily implemented using a separate 
coloring for each color in the multicoloring.

@@

%This function also demonstrates the benefit of being able to dynamically
%(re-)discover disjointness.  Safely running both instances of 
%{\tt calc\_new\_currents} in parallel depends on knowing that {\tt rw0} and
%{\tt rw1} in {\tt execute\_time\_steps} are disjoint.  Similarly, parallel
%execution of the instances of {\tt update\_voltage} requires knowing that
%{\tt rn0} is disjoint from {\tt rn1}.  In both cases, this knowledge was
%statically available in {\tt simulate\_circuit} and could have been captured
%in a region relationship, but it would have resulted in much more complicated
%code.
%Instead, an inexpensive dynamic disjointness check is performed by the Legion
%runtime, and the safety of parallel execution follows from the soundness of
%the type system.
The {\tt execute\_time\_steps} function also demonstrates the benefit of dynamically
being able to discover disjointness.  To run the two intances of 
{\tt calc\_new\_currents} in parallel requires knowing that {\tt rw0}
and {\tt rw1} are disjoint.  There is a similar requirement for parallel
execution of the two {\tt update\_voltage} calls with {\tt rn0} and {\tt rn1}.  In
both cases, this knowledge was statically available in {\tt simulate\_circuit}
and could have been captured in a region relationship at the cost
of much more complicated code.  Instead, an inexpensive dynamic disjointness
check by the Legion runtime will discover the parallelism.

@@

In addition to privileges, functions can also specify {\em coherence}
on regions.  Coherence specifies what updates the function may
see from other functions using the same region.  If not otherwise
specified, coherence defaults to {\em exclusive}, meaning the
function must appear to execute in program order.  Line 14 shows
an example of {\tt atomic}, a relaxed coherence mode requiring
that operations to {\tt rn} and {\tt rg} appear atomic relative
to other functions using those regions.  The most relaxed coherence
mode is {\tt simult}; simultaneous coherence allows concurrent access 
to the region by all functions with simultaneous coherence on the region.
